\documentclass{article}

\title{Using rectangular navigation meshes for pathfinding on square grid-based maps}
\author{Leonid Darovskikh \\
	\small Supervisor: Markus Sch\"utz
}
\date{\today}

\begin{document}

\maketitle

\section{Motivation \& Problem Statement}
Square grids are commonly used for pathfinding in many areas, such as mobile robot motion control, microchip routing and video games, in such genres as strategies, role-playing games, roguelikes and automation games. The most commonly used~\cite{cui2011based} algorithm is A*~\cite{hart1968formal}, which is optimal and efficient, but on large maps with many units performing pathfinding simultaneously, even it can have problems. Thus, programmers have been developing ways of accelerating it. One of such ways are navmeshes (navigation meshes)~\cite{snook2000simplified}, which subdivide the map into convex polygons and uses them as nodes of graph used to calculate the paths. Navmeshes commonly use triangles, since navigable areas can be represented as polygons, and every polygon can be triangulated, but in our experiments when the grid is represented as a polygon, subdividing it into rectangles yields far fewer polygons than with triangulation, even when using an suboptimal greedy algorithm. Fewer nodes should result in better performance, which makes rectangular navmeshes a promising approach for pathfinding on square grid-based maps. \\

However, there are several problems with this approach. Firstly, a rectangular navmesh would have to be generated, which can fortunately be done using a rectilinear polygon decomposition algorithm, which can solve this problem for a minimum amount of partitions~\cite{eppstein2009graph}. Secondly, our approach would have to be competitive with other high-performance pathfinding techniques designed specifically for grids, such as JPS~\cite{harabor2011online} and REA*~\cite{zhang2016rectangle}. These algorithms typically do not preprocess the grid, having to traverse it each time, which is where a navmesh can beat them, since it drastically reduces the node count. However, the fact that they require no preprocessing also means that they support dynamically changing terrain (e.g. buildings being created and serving as new obstacles) ``out of the box'', therefore, to stay competitive, we would have to implement some form of local navmesh repair to also support dynamic environments. Another way in which our technique could outcompete them is by supporting any-angle pathfinding, since they are usually limited by the directions of the grid neighborhood, although modifications of some of those algorithms exist that take care of that~\cite{luo2022improved}. Another feature that we would like our approach to provide is support for maps with non-uniform movement costs (i.e. map tiles can be comprised of different terrain with different cost associated with it), which becomes complicated when paired with any-angle pathfinding~\cite{mitchell1991weighted}. Lastly, a support for multi-agent pathfinding is desirable. \\

All of this together provides exciting possibilities for innovation.

\section{Expected Results}

The main result of this thesis will be an C++ library implementing the described pathfinding approach from start to finish. In order to test and debug it, an interactive application with a graphical user interface will also be implemented. It will support generating and editing the maps, calculating the paths and visualizing them. All of the code will be open source and available on Github. \\

To demonstrate that our approach yields an improvement in performance, the thesis will also include benchmarks against established algorithms such as A*, JPS and A* with string-pulling on a triangular navmesh in a variety of scenarios, such as different sizes, types and densities of maps. The recorded values will include run times and path lengths, as well as some other values, to be decided later~\cite{sturtevant2012benchmarks}. Finally, the benchmarks will be evaluated to determine the relationships between different scenarios and the performance gains of our technique. \\

We will also provide reference implementations of those established algorithms, if necessary (e.g. no implementation available or drastically different architecture, affecting the performance). 

\section{Methodology}
\begin{enumerate}
	\item \textbf{Literature review.} \\
	To begin with, an extensive survey of existing algorithms will be conducted in order to determine their strengths and weaknesses.
	\item \textbf{Formalization and algorithm design.} \\
	In this step we will decide on the algorithm's capabilities, design it and determine the map representation that it operates on. Also, we will choose map archetypes (e.g. maze, dungeon, cave, etc.) to conduct the benchmarking on as well as selecting the established algorithms to serve as a performance baseline.
	\item \textbf{Implementation.} \\
	Actually implementing the software, including the rectangular decomposition algorithm, pathfinding algorithm, and the benchmarking framework. If necessary - also the baseline algorithms and map generators. The programming language is C++.
	\item \textbf{Benchmarking.} \\
	Run our algorithm and the selected baseline algorithms and measure the runtimes and memory usage for different map types, sizes and densities.
	\item \textbf{Evaluation.} \\
	Using the measurements from the previous step, a statistical analysis will be conducted to determine how well our algorithm performs, and in which cases it offers an advantage over established methods.
\end{enumerate}

\section{State of the Art}

%\subsection{Pathfinding on a grid}
JPS(jump point search)~\cite{harabor2011online} is a modification of A* specifically for uniform grids with uniform movement cost. It achieves a speed-up by performing long straight "jumps" and only expanding the nodes where the jump ends (referred to as "\emph{jump points}"). In the tests conducted by the authors of the paper, it managed to be faster than A* by 25-30 times. The same authors improved the algorithm even further by precomputing jump points~\cite{harabor2014improving}, using bitwise operations to process whole blocks of tiles at once when scanning for jump points and improving the pruning rules.\\

RSR(rectangular symmetry reduction)~\cite{harabor2011path} decomposes the grid into obstacle-free rectangles, but instead of using them as navmesh nodes directly, it keeps the nodes making up the perimeter of rectangles and removes the one in the middle. Then it adds extra edges to perimeter nodes based on the grid connectivity (e.g. if the grid is 4-connected, only the nodes on the opposite sides are connected). This improves the performance of A*, but perimeter nodes still have to traversed normally, to deal with that, the authors suggest pruning the perimeter nodes as well.

Both JSP and RSR are based on the idea of symmetry, which in this case means existence of paths between two nodes that are different, but have the same length and are therefore essentially equivalent. But search algorithms perform redundant work by evaluating them, thus reducing symmetry leads to improved performance. \\

HPA*(hierarchical pathfinding A*)~\cite{botea2004near} subdivides the map \emph{clusters}: equally sized disjoint squares. It then detects \emph{entrances}, obstacle-free segments along the borders of adjacent clusters. After that, it builds an abstract representation of the map as a graph with entrances as nodes and edges connecting them if they belong to the same cluster and there exists a path between them or if they belong to neighbouring clusters and are separated by the border. This graph is used to calculate the abstract path which is then refined by finding paths between every two consecutive entrances within clusters using the actual map. This method can achieve a 10x speed-up, but requires preprocessing and tends to generate suboptimal paths. \\

There are also derivative algorithms of HPA*, such as HAA*(hierarchical annotated A*)~\cite{harabor2008hierarchical}, which takes agent size and traversal capabilities(e.g. can traverse ground but cannot traverse swamp) into account by computing clearance maps for every agent capability, annotating the edges between clusters with maximum clearance and required capabilities. Another algorithm is DHPA*(Dynamic HPA*)~\cite{kring2010dhpa} which caches paths from every tile of a cluster to abstract nodes in it and then uses those paths for low-level pathfinding, which allows it to achieve a 10 times speed-up compared to regular HPA*. And if the map is altered, only the caches for the affected clusters need to be recalculated. \\

REA*(rectangle expansion A*)~\cite{zhang2016rectangle} uses maps with uniform movement cost, similarly to JPS, processing them as obstacle-free rectangles and only using their boundaries as search nodes. Any point on a boundary can be chosen as a successor, so the path is not restricted to the grid's cardinal directions and can be shorter than grid-optimal, but it is not truly any-angle. In the experiments conducted by the paper's authors, REA* showed that in certain cases it can outperform A* by 25 times, which can rival JPS.

%Polyanya

%\subsection{Navigation meshes}
%NEOGEN~\cite{oliva2013neogen}

%ECM(explicit corridor maps)~\cite{geraerts2010planning} is a navmesh represented as a graph whose edges are arcs of the medial axis of the walkable space and vertices - points where they connect, also annotated with distances to the closest obstacle, allowing the the ECM to take clearances into account.

\section{Relevance to the Visual Computing Curriculum}
This thesis deals with many areas of computer science: algorithms and data structures, robotics and motion planning, graph theory, computational geometry. Not all of them are covered by the visual computing master programme, but the following courses are the most relevant ones:

\begin{itemize}	
	\item 186.122 Algorithmic Geometry
	\item 186.814 Algorithmics
	\item 104.283 Discrete Mathematics for Computer Science
	\item 192.162 Heuristic Optimization Techniques
\end{itemize}

%\nocite{*}
\bibliographystyle{plain}
\bibliography{bibliography}
\end{document}